---
title: "Assignment 2"
author: "Tejasvini Mavuleti"
date: "4/9/2022"
output:
  word_document: default
  pdf_document: default
---

Setting working directory
```{r}
getwd()
setwd("C:/Users/mavul/OneDrive/Documents")
```

# 1) Running the code to create two variables X and Y

```{r}
set.seed(2017) 
X=runif(100)*10 
Y=X*4+3.45 
Y=rnorm(100)*0.29*Y+Y
```

## a) Ploting Y against X

```{r}
plot(X,Y, xlab = "X (units)", ylab = "Y (units)", main = "Plot of Y vs X ", col = "green")
```
From the graph shown above there is a positive linear trend between X and Y variables
 
## b) Constructing a simple linear model of Y based on X

```{r}
Model <- lm(Y~X)
Model$coefficients
```

The formula to explain Y based on X from our linear model is: Y = 3.6108*X + 4.4655

## c) The relation of the coefficient of determination and coefficient correlation of X and Y

```{r}
summary(Model)
```
The R2 is 0.6517, that means 65% of the variability of Y is captured by X


# 2) Including the ‘mtcars’ dataset included in R distribution

```{r}
# shows first 6 rows
head(mtcars)
```

## a) Constructing simple linear models using mtcars data 

```{r}
# Creating a linear model for weight vs horsepower and displays a plot of the points

Model2 = lm(hp~wt, data = mtcars)
summary(Model2)
plot(mtcars$wt,mtcars$hp, xlab = "Weight (wt)", ylab = "Horsepower (hp)", main = "Plot of Weight vs Horsepower", col = "red")
```
From this linear model we can see that weight results in a model that accounts for 43.39% of the variation in horsepower

```{r}
# Creating a linear model for mpg vs horsepower and displays a plot of the points
Model3 = lm(hp~mpg, data = mtcars)
summary(Model3)
plot(mtcars$mpg,mtcars$hp, xlab = "Miles per Gallon (mpg)", ylab = "Horsepower (hp)", main = "Plot of Miles Per Gallon vs Horsepower", col = "blue")
```

From this linear model the fuel efficiency results in a model that accounts for 60.24% of the variation of the horsepower. Therefore, the fuel efficiency (mpg) is considered statistically significant in this model.


## b) Building a model that uses the number of cylinders (cyl) and the mile per gallon (mpg) values of a car to predict the car Horse Power (hp)

```{r}
# Shows which variables are factor or numeric
str(mtcars)
# Converting cylinder into a factor
mtcars$cyl = as.factor(mtcars$cyl)
str(mtcars)
```

A multiple regression model is used in this case to build a model that represents horsepower as a result of cylinders and miles per gallon

```{r}
Model5 = lm(hp~cyl+mpg, data = mtcars)
summary(Model5)
# Predict the estimated horse power of a car with 4 cylinders and 22 mpg
predict(Model5, data.frame(mpg = c(22), cyl = c("4")))
```

The estimated Horse Power of a car with 4 calendar and mpg of 22 is 98.15%


# 3) Using BostonHousing dataset

```{r}
library(mlbench)
data(BostonHousing)
```

## a) Building a model to estimate the median value of owner-occupied homes 

```{r}
str(BostonHousing)
```

```{r}
# Creating a linear model for median value based on crim, zn, ptratio, and chas.
Model6 = lm(medv~crim+zn+ptratio+chas, data = BostonHousing)
summary(Model6)
```
The R2 value in this model (crim, zn, ptratio, and chas) is 35.99% of the variability in median housing value. This is a weak model in terms of accuracy and can be improved by adding more variables into the model.

## b) Using the estimated coefficient
I) Based on the coefficients, the resulting formula from our model is:

medv = 49.91868 - 0.26018*crim + 0.07073*zn - 1.49367*ptratio + 4.58393*chas1

Therefore, if the only difference between two houses is that one borders the Chas River, then we focus on the chas variable coefficient. The house that borders the river would be $4,583.93 more than the one that does not.

4.58393 (coeff of chas) * 1 (value of chas) * 1000 (medv in $1,000 units) = $4,583.93

II)	Based on the coefficients, the resulting formula from our model is:

medv = 49.91868 - 0.26018*crim + 0.07073*zn - 1.49367*ptratio + 4.58393*chas1

Therefore, if the only difference between two houses is the pupil-teacher ratio, then we focus on the ptratio variable coefficient. As a result, the house with the smaller pupil-teacher ratio value is more expensive, because the coefficient is found to be negative in our model. The difference in values between the houses is

-1.49367 (coeff of ptratio) * 0.03 (difference between ptratio values) * 1000 (medv in $1,000 units) = $44.81

Therefore, the house with the lower pupil-teacher ratio is $44.81 more expensive based on our model

## c)	Which of the variables are statistically important?

Based on the model constructed from these variables, all of the variables (crim, zn, ptratio, and chas) were found to be statistically significant. 
This is true because all of the p-values calculated from our model at below the 0.05 threshold value for significance.

## d)	Using the anova analysis and determine the order of importance of these four variables 

```{r}
# Returns the ANOVA results for the model used in this problem
anova(Model6)
```

Based on the ANOVA values returned, the order of importance of these variables are:

1. "crim" - accounts for 15.08% of variability in the model
2. "ptratio" - accounts for 11.02% of variability in the model
3. "zn" - accounts for 8.32% of variability in the model
4. "chas" - accounts for 1.56% of variability in the model

The residuals in this model still account for 64.01% of variability in the model.
There is still a lot of room for improvement in the accuracy of this model.
 
